### 2016-03-11 docs/cache/appendix_mathsupport.Rmd
adds online supporting information document

>>>>

Equations A# and A# show that synchrony population biomasses and growth rates emerge from complex interactions among species' intrinsic growth rates, interspecific interactions, environmental stochasticity, and demographic stochasticity.
Given these complexities, it is impossible to determine expected effects of different parameters in a multi-species case.
Thus, we analyze a simplified case where interspecific interactions are zero.
The variance-covariance matrix of population biomasses at steady state then becomes

\begin{equation}
c_{ij} = \frac{B_{ij}}{1-a_i a_j} = \frac{N_{i}^* N_{j}^* \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \sqrt{N_{i}^* N_{j}^*} \sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})}{1 - (1-r_{mi})(1-r_{mj})}
\end{equation}

and the variance-covariance matrix of population growth rates becomes

\begin{equation}
R_{ij} = \frac{r_{mi}r_{mj}}{K_i K_j} C_{ij} + \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})}{\sqrt{N_{i}^* N_{j}^*}},
\end{equation}

which simplifies to

\begin{equation}
R_{ij} = \frac{1}{1- \frac{r_{mi}r_{mj}}{r_{mi}+r_{mj}}} \left( \sigma_{ei} \sigma_{ej} \text{cov}(\mu_{ei},\mu_{ej}) + \frac{\sigma_{di}\sigma_{dj} \text{cov}(\mu_{di},\mu_{dj})}{\sqrt{N_{i}^* N_{j}^*}} \right)
\end{equation}

\noindent This is still not a general result, however, because synchrony in population sizes and synchrony in growth rates use weighted factors of the environmental variances and covariances with species-specific parameters.
So, we further assume that, along with interspecific interactions being zero, all species have identical growth rates, environmental stochasticity is absent, and all species have identical demographic variance.
This represents a theoretical limiting case where the community consists of identical species coexisting in a constant environment where only demographic stochasticity causes temporal fluctuations.
Under such conditions, synchrony of population biomasses is

\begin{equation}
\phi_N = \frac{1}{\left(\sum_i p_i^{1/2} \right)^2}
\end{equation}

\noindent and synchrony of growth rates is

\begin{equation}
\phi_N = \frac{\sum_i p_i^{-1}}{\left(\sum_i p_i^{-1/2} \right)^2}
\end{equation}

\noindent where $p_i$ is the average frequency of species *i*, $p_i = N_i/N_T$.
When all species have identical abundances and $p_i = 1/S$, where *S* is species richness, the both synchrony values equal 1/S [@Loreau2008a].

### 2016-03-11 docs/ms_communitySynchrony.Rmd
adds online supporting information document

To derive analytical predictions we...CLAIRE'S TEXT ON DERIVING THE PREDICTION...

@@

Qualitative predictions here...?

@@

#### Fitting statistical models
Vital rate regressions are the building blocks of our dynamic models: an integral projection model (IPM) and an individual-based model (IBM).
We followed the approach of @Chu2015 to fit statistical models for survival, growth, and recruitment.
We modeled survival probability of each genet as function of genet size, temporal variation among years, permanent spatial variation among groups of quadrats, and local neighborhood crowding from conspecific and heterospecific genets.
Regression coefficients for the effect of crowding by each species can be considered a matrix of interaction coefficients whose diagonals represent intraspecific interations and whose off-diagonals represent interspecific interactions [@Adler2010]. These interaction coefficients can take positive (facilitative) or negative (competitive) values.
We modeled growth as the change in size of a genet from one year to the next, which depends on the same factors as the survival model.
We fit the survival and growth regressions using INLA [@Rue2014], a statistical package for fitting generalized linear mixed effects models via approximate Bayesian inference [@Rue2009], in R [@R2013].
Genet size and crowding were treated as fixed effects; spatial (quadrat groupings) and temporal (interannual) variation were treated as random effects on the intercept.

@@

The IPM uses the same vital rate regressions as the IBM [@Rees2009; @Rees2014], but it is spatially-implicit and does not include demographic stochasticity.
Following @Chu2015, we use a mean field approximation that captures the essential features of spatial patterning to define the crowding index at each time step.
Temporal variation is included in exactly the same way as for the IBM.
For full details on the IPMs we use, see @Chu2015.

@@

### Results
Synchrony of species' per capita growth rates at our study sites ranged from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to 0.92 (Table 2).
Synchrony tended to be higher in communities with few species (Arizona and New Mexico) and/or with relatively high temporal variability in percent cover (e.g., Montana).
Synchrony was lowest in Idaho, the only data set that includes two life forms: a shrub and three perennial grasses (Table 1).
For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig. X1).
IBM-simulated synchrony is consistently, but only slightly, lower than IPM-simulated synchrony (Fig. X1), likely due to the desynchronizing effect of demographic stochasticity.

@@

Simulation experiments revealed that removing environmental fluctuations has the largest impact on synchrony, leading to a reduction in synchrony of species growth rates in four out five communities (Fig. 2).
Removing environmental fluctuations ("No E.S" simulations) decreased synchrony by 33% in Arizona, 48% in Kansas, 39% in Montana, and 40% in New Mexico.
Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2), but the effect was small (9% increase).
Species interactions are weak in these communities [Table SX and @Chu2015], and removing interspecific competition had little effect on synchrony (Fig. 2; "No Comp." simulations).
Removing competition caused, at most, a 5% change in synchrony (Table SX).
Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig. 2), with an average 6% increase [IS THIS FOR THE SMALLEST SIMULATED PLOT SIZE?].
The desynchronizing effect of demographic stochasticity modestly counteracted the synchronizing force of the environment, but not enough to lower synchrony to the level observed when only demographic stochasticity is operating (Fig. 3).
For all five communities, the synchrony of species' growth rates in monoculture was strongly associated with species synchrony in polyculture (rank correlation = 1; Fig. 3).
Results for synchrony of percent cover are qualitatively similar (Figs. SX-SX) but more variable, as expected by theory [@Loreau2008a].

@@

### Simulations support theoretical predictions
Our theoretical predictions were derived from a simple model of population dynamics and required several simplifying assumptions, raising questions about their relevance to natural communities.
However, the theoretical predictions were strongly supported by simulations of population models fit to long-term data from natural plant communities (Table 3).
Such a close match between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony.
The theoretical population dynamical model we used here also successfully predicted ecosystem stability in several biodiversity-ecosystem function experiments [@DeMazancourt2013a].
The accumulating evidence from both experimental and observational studies demonstrates the utility of linking population dynamics to community and ecosystem level properties [@Loreau2010]. [THIS LAST SENTENCE SUCKS]

@@

In the Idaho community, removing environmental fluctuations did not cause a large decrease in synchrony.
However, that result appears to be an artifact. Removing environmental variation results in a negative invasion growth rate for *A. tripartita*. Although we only analyzed IBM runs in which *A. tripartita* had not yet gone extinct, it was at much lower abundance than in the other simulation runs.
When we removed the shrub (*Artemisia tripartita*) from all simulations, the Idaho results conformed with results from all other sites: removing environmental stochasticity caused a significant reduction in species synchrony (Fig. SX).
Our main results for Idaho (Fig. 2), with *A. tripartita* included, demonstrate how the processes that determine species synchrony interact in complex ways.
*A. tripartita* has a facilitative effect on each grass species across all vital rates, except for a small competitive effect on *H. comata*'s survival probability (Tables SX, SX, SX).
At the same time, all the perennial grasses have negative effects on each other for each vital rate (Tables SX, SX, SX).
We know synchrony is affected by interspecific competition [@Loreau2008a], but how facilitative effects manifest themselves is unknown.
The interaction of facilitation and competition is clearly capable of having a large effect on species synchrony, and future theoretical efforts should aim to include a wider range of species interactions.

@@

Our results do not contradict the idea that competition can lead to compensatory dynamics, but they do highlight the fact that interspecific competition must be relatively strong to influence species synchrony.
The communities we analyzed are composed of species with very little niche overlap [@Chu2015] and weak interspecific interactions (Tables X1-X10).
Mechanistic consumer-resource models [@Lehman2000] and phenomenological Lotka-Volterra models [@Lehman2000; @Loreau2013] both confirm that the effect of competition on species synchrony diminishes as niche overlap increases.
In that sense, our results are not surprising: interspecific interactions are weak, so of course removing them does not affect synchrony.

>>>>

To derive analytical predictions we...CLAIRE'S TEXT ON DERIVING THE PREDICTION... (see Online Supporting Information).

@@

@@

#### Fitting statistical models
Vital rate regressions are the building blocks of our dynamic models: an integral projection model (IPM) and an individual-based model (IBM).
We followed the approach of @Chu2015 to fit statistical models for survival, growth, and recruitment (see Online Supporting Information for full details).
We modeled survival probability of each genet as function of genet size, temporal variation among years, permanent spatial variation among groups of quadrats, and local neighborhood crowding from conspecific and heterospecific genets.
Regression coefficients for the effect of crowding by each species can be considered a matrix of interaction coefficients whose diagonals represent intraspecific interations and whose off-diagonals represent interspecific interactions [@Adler2010]. These interaction coefficients can take positive (facilitative) or negative (competitive) values.
We modeled growth as the change in size of a genet from one year to the next, which depends on the same factors as the survival model.
We fit the survival and growth regressions using INLA [@Rue2014], a statistical package for fitting generalized linear mixed effects models via approximate Bayesian inference [@Rue2009], in R [@R2013].
Genet size and crowding were treated as fixed effects; spatial (quadrat groupings) and temporal (interannual) variation were treated as random effects on the intercept.

@@

The IPM uses the same vital rate regressions as the IBM [@Rees2009; @Rees2014], but it is spatially-implicit and does not include demographic stochasticity.
Following @Chu2015, we use a mean field approximation that captures the essential features of spatial patterning to define the crowding index at each time step (Supporting Online Information).
Temporal variation is included in exactly the same way as for the IBM.
For full details on the IPMs we use, see @Chu2015.

@@

### Results
Synchrony of species' per capita growth rates at our study sites ranged from 0.36 to 0.89 and synchrony of percent cover ranged from 0.15 to 0.92 (Table 2).
Synchrony tended to be higher in communities with few species (Arizona and New Mexico) and/or with relatively high temporal variability in percent cover (e.g., Montana).
Synchrony was lowest in Idaho, the only data set that includes two life forms: a shrub and three perennial grasses (Table 1).
For all five communities, species synchrony from IPM and IBM simulations closely approximated observed synchrony (Fig. S1).
IBM-simulated synchrony is consistently, but only slightly, lower than IPM-simulated synchrony (Fig. S1), likely due to the desynchronizing effect of demographic stochasticity.

@@

Simulation experiments revealed that removing environmental fluctuations has the largest impact on synchrony, leading to a reduction in synchrony of species growth rates in four out five communities (Fig. 2).
Removing environmental fluctuations ("No E.S" simulations) decreased synchrony by 33% in Arizona, 48% in Kansas, 39% in Montana, and 40% in New Mexico.
Only in Idaho did removing environmental fluctuations cause an increase in synchrony (Fig. 2), but the effect was small (9% increase; Table S2).
Species interactions are weak in these communities [Table S3 and @Chu2015], and removing interspecific competition had little effect on synchrony (Fig. 2; "No Comp." simulations).
Removing competition caused, at most, a 5% change in synchrony (Table S2).
Removing demographic stochasticity ("No D.S." simulations) caused synchrony to increase slightly in all communities (Fig. 2), with an average 6% increase over synchrony from IBM simulations on a five meter area.
The desynchronizing effect of demographic stochasticity modestly counteracted the synchronizing force of the environment, but not enough to lower synchrony to the level observed when only demographic stochasticity is operating (Fig. 3).
For all five communities, the synchrony of species' growth rates in monoculture was strongly associated with species synchrony in polyculture (rank correlation = 1; Fig. 3).
Results for synchrony of percent cover are qualitatively similar (Figs. S2-S3) but more variable, as expected by theory [@Loreau2008a].

@@

### Simulations support theoretical predictions
Our theoretical predictions were derived from a simple model of population dynamics and required several simplifying assumptions, raising questions about their relevance to natural communities.
However, the theoretical predictions were strongly supported by simulations of population models fit to long-term data from natural plant communities (Table 3).
Such a close match between our analytical predictions and the simulation results should inspire confidence in the ability of simple models to inform our understanding of species synchrony.
The theoretical population dynamical model we used here also successfully predicted ecosystem stability in several biodiversity-ecosystem function experiments [@DeMazancourt2013a].

@@

In the Idaho community, removing environmental fluctuations did not cause a large decrease in synchrony.
However, that result appears to be an artifact. Removing environmental variation results in a negative invasion growth rate for *A. tripartita*. Although we only analyzed IBM runs in which *A. tripartita* had not yet gone extinct, it was at much lower abundance than in the other simulation runs.
When we removed *A. tripartita* from all simulations, the Idaho results conformed with results from all other sites: removing environmental stochasticity caused a significant reduction in species synchrony (Fig. S4).
Our main results for Idaho (Fig. 2), with *A. tripartita* included, demonstrate how the processes that determine species synchrony interact in complex ways.
*A. tripartita* has a facilitative effect on each grass species across all vital rates, except for a small competitive effect on *H. comata*'s survival probability (Tables S4, S9, S14).
At the same time, all the perennial grasses have negative effects on each other for each vital rate (Tables S4, S9, S14).
We know synchrony is affected by interspecific competition [@Loreau2008a], but how facilitative effects manifest themselves is unknown.
The interaction of facilitation and competition is clearly capable of having a large effect on species synchrony, and future theoretical efforts should aim to include a wider range of species interactions.

@@

Our results do not contradict the idea that competition can lead to compensatory dynamics, but they do highlight the fact that interspecific competition must be relatively strong to influence species synchrony.
The communities we analyzed are composed of species with very little niche overlap [@Chu2015] and weak interspecific interactions (Tables S1, S3-S17).
Mechanistic consumer-resource models [@Lehman2000] and phenomenological Lotka-Volterra models [@Lehman2000; @Loreau2013] both confirm that the effect of competition on species synchrony diminishes as niche overlap increases.
In that sense, our results are not surprising: interspecific interactions are weak, so of course removing them does not affect synchrony.
